Functory: Distributed Computing for the Common Man\thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2010-10-11 doc/report/main.tex
*** empty log message ***

In the network configuration, we deliberately choose to have two
different programs for master and workers, using the \of{Network.Poly}
module (actually, a single source code with a command line option).
This has to be done since the worker code makes use of global
variables, which prevents us from using
\of{Network.Same}, as described in Section~\ref{sec:API}. 
Alternately, we could have refactored the worker
code to avoid global variables.

@@

@@

\subsubsection{Three Network Implementations} % TODO: change title

As mentioned in Section~\ref{sec:API}, the \of{Network} module
actually provides three different implementations, according to three
different execution scenarios (namely, \of{Same}, \of{Poly} and
\of{Mono}). There is
actually a single, generic implementation, which is then instantiated
in three different ways, corresponding to the three scenarios.  The
generic implementation assumes that only strings are being transmitted
over the network, as evident from the \of{Assign} and \of{Completed}
messages in Section~\ref{sec:protocol}. Depending on the
implementation, these strings could be marshaled closures, marshaled
values or simply user strings.

>>>>

In the network configuration, we deliberately choose to have two
different programs for master and workers, using the \of{Network.Poly}
module (actually, a single source code with a command line option).
This has to be done since the worker code makes use of global
variables, which prevents us from using
\of{Network.Same} (as explained below in Section~\ref{sec:implem}). 
Alternately, we could have refactored the worker
code to avoid global variables.

@@

\subsubsection{Network Sub-modules} % TODO: change title

As mentioned in Section~\ref{sec:API}, the \of{Network} module
actually provides three different implementations, according to three
different execution scenarios. There are provided in three
sub-modules, as described below.

\paragraph{\of{Same}.} This module is used when master and workers are
running the same executable. The master and workers have to be
differentiated in some manner. We use an environment variable
\texttt{WORKER} for this purpose. When set, it indicates that the
executable acts as a worker. At run-time, a worker immediately 
enters a loop waiting for tasks from the master, without even getting
into the user code. 
As explained in Section~\ref{sec:API}, 
the master function has the following signature.
\begin{ocaml}
 val compute : 
      worker:('a -> 'b) -> 
      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit  
\end{ocaml}
The master uses marshaling to send both a closure of type 
\of{'a -> 'b} and a task of type \of{'a} to the worker. The resulting strings
are passed as argument \of{f} and \of{x} in message
\of{Assign}. Similarly, the worker uses marshaling to send back the
result of the computation of type \of{'b}, which is the argument
\of{s} in message \of{Completed}.

Though the ability to run the same executable helps a lot in deploying
the program in different machines, it comes at a small price. Since
the worker is not getting into the user code, closures which are
transmitted from the master can not refer to global variables in the
user code. Indeed, the initialization code for these global variables
is never reached on the worker side. For instance, the Mandelbrot set
example could be written as follows:
\begin{ocaml}
  let max_iterations = 200
  let worker (xmi, xma, ymi, yma, w, h) =
    ... draw sub-image using max_iterations ...
\end{ocaml}
That is, the global function \of{worker} makes use of the global
variable \of{max_iterations}. The worker gets the function to compute
from the master, namely the closure corresponding to function
\of{worker} in that case, but on the worker side the initialization of
\of{max_iterations} is never executed.

One obvious solution is  not to use global variables in the
worker code. This is not always possible, though. 
To overcome this, the \of{Same} sub-module also provides a
\of{Worker.compute} function to start the worker loop manually from
the user code. This way, it can be started at any point, in particular
after the initialization of the required global variables.
Master and worker are still running the same executable, but are 
distinguished using a user-defined way (command-line argument,
environment variable, etc.).

\bigskip
There are situations where it is not possible
to run the same executable for master and workers.
For instance, architectures or operating systems could be different
across the network.
For that reason, the \of{Network} module provides two other
implementations.

\paragraph{\of{Poly}.} When master and workers are compiled with the
same version of \Ocaml, we can no longer marshal closures but we can
still marshal polymorphic values. Indeed, an interesting property of
marshaling in \Ocaml\ is to be fully architecture-independent, as long
as a single version of \Ocaml\ is used. It is worth pointing out that
absence of marshaled closures
now enables the use of two different programs for master and workers. This
is not mandatory, though, since master and workers could still be
distinguished at run-time as in the previous case.

On the worker side, the main loop is started manually using
\of{Worker.compute}. The computation to be performed on each task is
given as an argument to this function. It thus looks as follows:
\begin{ocaml}
  Worker.compute : ('a -> 'b) -> unit -> unit
\end{ocaml}
On the master side, the \of{compute} function is simpler than in the
previous case, as it has one argument less, and thus has the following
signature. 
\begin{ocaml}
  Master.compute : master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit 
\end{ocaml}
For realistic applications, where master and workers are completely
different programs, possibly written by different teams, this is the
module of choice in our library, since it can still pass polymorphic
values over the network. The issues of marshaling are automatically
taken care of by \Ocaml\ run-time.

% TODO : explain the new derived API

\paragraph{\of{Mono}.}

% TODO

% \bigskip

% There is
% actually a single, generic implementation, which is then instantiated
% in three different ways, corresponding to the three scenarios.  The
% generic implementation assumes that only strings are being transmitted
% over the network, as evident from the \of{Assign} and \of{Completed}
% messages in Section~\ref{sec:protocol}. Depending on the
% implementation, these strings could be marshaled closures, marshaled
% values or simply user strings.

@@
